### 2021-03-29 proposal/proposal.Rmd
outlined project proposal

---
output: 
  stevetemplates::article:
    fig_caption: true
#bibliography: master.bib
urlcolor: blue
biblio-style: apsr
title: "Using Statistical Methods to Predict the Outcome of Basketball Games"
author:
- name: Sabrina Ball
  affiliation:
- name: Joseph Despres
  affiliation:
abstract: "Quick Summary"
geometry: margin=1in
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
endnote: no
---

```{r eval=FALSE, include=FALSE}
op <- par(mar = rep(0, 4))
plot(imager::load.image("https://i.imgur.com/f0KNGaD.png"))
par(op)
```

```{r, include=FALSE}
source(here::here("R", "explatoryAnalysis.R"))
```

# 1. Data

Our data comes from three different data sources Kaggle, NCAA, and a March Madness bracket. First, a Kaggle^[https://www.kaggle.com/c/ncaam-march-mania-2021/data] machine learning challenge that has a comprehensive records of every US division I basketball game played from 2003 to 2020. Second, the NCAA^[http://stats.ncaa.org/rankings/change_sport_year_div] team statistics. Third, the tournament bracket in a nice .xls format we obtained from a blog^[https://plexkits.com/march-madness-bracket/].

## Variables

Our response is the outcome win/loss as a function of team statistics. 
win 
`r #TODO: fill this out maybe make this a list if its too big`

|  Variable name     |  Description  |   Type  |
| ------------------ | ------------- | ------- |
| win                |               |         |
| fg_percent         |               |         |
| opposingfg_percent |               |         |
| ft_percent         |               |         |
| opposingft_percent |               |         |
| rpg                |               |         |
| opposingrpg        |               |         |
| st                 |               |         |
| opposingst         |               |         |
| to                 |               |         |
| opposingto         |               |         |
| opposingbkpg       |               |         |
| bkpg               |               |         |

# 2. Questions

Can we use statistical methods to predict outcomes of the March Madness basketball games better than chance? Better than seed? Better than betting markets?

# 3. Learning

We would like to learn how to make predictions with logistic regression. We would like to come up with a good way to evaluate our predictions. This will be dificult since the outcome is binary so we would need to weight our predictions somehow and index our socres.

# 4. Methods

Logistic regression with a response prediction function.

# 5. Issues

The seeds are likely derrived from the predictors we are using. Multicolinearity, will be something to watchout for because better teams probally have better statistics.