### 2021-04-08 slides/slides.Rmd
adding slides

---
title: "March Madness Basketball Turniment"
author: "Sabrina Ball and Joe Despres"
date: "updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

background-image: url(https://upload.wikimedia.org/wikipedia/en/2/28/March_Madness_logo.svg)

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

---

https://github.com/EvaMaeRey/flipbookr

---