### 2021-04-18 paper/formula.rmd
worked on paper

---
title: ""
subtitle: ""
author: ""
institute: ""
date: ""
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r eval=FALSE, echo=FALSE}
equatiomatic::extract_eq(model = logistic_fit, 
                         use_coefs = FALSE, 
                         show_distribution = FALSE,
                         intercept = "beta")
# results='asis' in code chunk to render as LaTex
```

$$
\begin{align}
\beta_{0} + \beta_{1}(\operatorname{x3fg}) + \beta_{2}(\operatorname{opposingx3fg}) + \beta_{3}(\operatorname{fg\_percent}) + \beta_{4}(\operatorname{opposingfg\_percent}) + \beta_{5}(\operatorname{ft\_percent}) + \beta_{6}(\operatorname{opposingft\_percent}) + \beta_{7}(\operatorname{rpg}) + \beta_{8}(\operatorname{opposingrpg}) + \beta_{9}(\operatorname{st}) + \beta_{10}(\operatorname{opposingst}) + \beta_{11}(\operatorname{to}) + \beta_{12}(\operatorname{opposingto}) + \beta_{13}(\operatorname{opposingbkpg}) + \beta_{14}(\operatorname{bkpg})
\begin{align}
$$

---
$$
\begin{align}
\operatorname{polrights} = \beta_{0} + \beta_{1}(\operatorname{corruption\_perception\_index\_cpi}) +\\ \beta_{2}(\operatorname{edu\_exp\_gdp\_per\_person}) +\\ \beta_{3}(\operatorname{life\_expectancy\_years}) +\\ \beta_{4}(\operatorname{military\_spending\_pct\_of\_gdp}) +\\ \beta_{5}(\operatorname{murder\_per\_mil\_people}) + \epsilon
\end{align}
$$